---
title: "How to reconcile forecasts with fable.bayesRecon"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{How to reconcile forecasts with fable.bayesRecon}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoded{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(tibble.width = Inf, knitr.kable.NA = "")

# Helper to render fabletools tables as HTML-safe kable
kable <- function(x, ...) {
  x |>
    dplyr::mutate(dplyr::across(
      where(~ inherits(.x, "agg_vec")),
      ~ gsub("<", "&lt;", gsub(">", "&gt;", as.character(.x)))
    )) |>
    knitr::kable(...)
}
```


## Introduction

This vignette shows how to use the reconciliation methods implemented in the `fable.bayesRecon` package.
In particular, we will apply t-Rec method [@carrara2025] to the `swiss_tourism` dataset, which contains monthly counts of tourists in various cantons of Switzerland. 

The method t-Rec is implemented in the function `bayesRecon_t()`, which can be used within the `reconcile()` function of `fabletools` to specify the reconciliation strategy.

First, we load the required packages:

```{r libraries}
library(bayesRecon)
library(dplyr)
library(tsibble)
library(fable)
library(fabletools)
library(fable.bayesRecon)
library(ggplot2)
```


## Data preparation

Let's import the data in the tsibble format. To do so, we follow these steps:

1. Load the `swiss_tourism` dataset from the `bayesRecon` package. This is a `mts` (multivariate time series) object. 
2. Take the `ts` component of the dataset, which contains the time series data.
3. Drop the first column, which contains the aggregated data for all cantons; it will be reconstructed later.
4. Convert it to a tsibble format using `as_tsibble()`.
5. Optionally, rename the columns for clarity with `rename()`.
6. Aggregate the data by canton using `aggregate_key()`.


```{r data}

data <- swiss_tourism$ts[, -1, drop = FALSE] |>
  as_tsibble() |>
  rename(Month = index, Canton = key, Tourists = value) |>
  aggregate_key(Canton, Tourists = sum(Tourists))

data |> head(5) |> kable()
```

The time series obtained in this way are stacked vertically, with a column standing for the time index (`Month`),  one for the grouping variable (`Canton`), and one for the values(`Tourists`).
However, `fabletools` allows to handle more complex hierarchies with more grouping variables.
Aggregated time series have their `Canton` value set to `<aggregated>`.


## Models fitting

We start by removing from each time series the last 12 observations, which will be used as test set, using the `filter()` function and the `yearmonth()` format to specify the cutoff date. This allows us to evaluate the performance of the reconciliation method on the holdout set.

Then, we fit a base forecasting model to the data, then we will reconcile the results. For base forecasts, we use an ETS model with additive trend and seasonality, with the `ETS()` function.

```{r fit}
fit <- data |>
  filter(Month < yearmonth("2024 Feb")) |>                  # Remove test set
  model(base = ETS(Tourists ~ trend("A") + season("A"))) |> # fit ETS model
  reconcile(t = bayesRecon_t(base),
            mint = min_trace(base))                         # Reconcile with t-Rec

fit |> head(5) |> kable()
```
**NOTA: vogliamo modificare qualocsa in modo tale che la classe sia qualcosa come `reconc_t`?**
**Direi che va bene com'è adesso: l'oggetto interno dentro fit è una classe `lst_bayesRecon_t` che e' l'equivalente di `lst_mint_mdl` per mint. quando visualizzi entrambi gli oggetti fa vedere la stessa cosa.**


During this step, base forecasts are fitted, but predictions are not generated yet.
Thus, the specification of the reconciliation strategy here is only used to set up the model.
The actual reconciliation will be performed when we produce forecasts.

## Reconciliation

Given the fitted models, we can now make one-year ahead predictions.
This is done applying the `forecast()` method to the fitted model.
The forecast horizon can either be specified as a number of periods (e.g., `h = 12` for 12 months) or as a date (e.g., `h = "1 year"`).

```{r forecast}
fc <- fit |>
  forecast(h = "1 year")

fc |> head(5) |> kable()
```

The forecasts object contains both base and reconciled forecasts. 
Their category, `base` or `t`, is specified in the `.model` column.
As in the fitted models, `Canton` and `Month` are respectively the aggregating variable and the time index.
The predicted varible, `Tourists`, contains object from the `distributional` package.
Point forecasts are returned as well, in the `.mean` column, which contains the mean of the predictive distribution.

## Summary of results

To compute the mean of each accuracy measure grouped by model:

```{r accuracy}
results <- fc |>
  accuracy(data, measures = list(RMSSE = RMSSE, MASE = MASE, CRPS = CRPS)) |>
  group_by(.model) |>
  summarise(RMSSE = mean(RMSSE), MASE  = mean(MASE), CRPS  = mean(CRPS))

results |> kable(digits = 3)
```



## Visualisation

We can visualize the forecasts using `autoplot()`. Below we plot the aggregated series and two individual cantons, showing both base and reconciled forecasts against the actual data.

```{r plot-aggregated, fig.width=7, fig.height=4}
# Aggregated series
fc |>
  filter(is_aggregated(Canton)) |>
  autoplot(data) +
  labs(title = "Aggregated Swiss Tourism", y = "Tourists", x = NULL) +
  theme_minimal()
```

```{r plot-canton1, fig.width=7, fig.height=4}
# First canton
cantons <- data |>
  filter(!is_aggregated(Canton)) |>
  pull(Canton) |>
  unique()

fc |>
  filter(Canton == cantons[1]) |>
  autoplot(data) +
  labs(title = paste("Canton:", cantons[1]), y = "Tourists", x = NULL) +
  theme_minimal()
```


```{r plot-canton2, fig.width=7, fig.height=4}
# Second canton
ggsave('plt.png', fc |>
  filter(Canton == "Luzern") |>
  autoplot(data |> filter(Month >= yearmonth("2022 Jan"),
                                 Canton == "Luzern"),
           level = 80) +
  labs(title = paste("Canton Luzern"), y = "Tourists", x = NULL) +
  theme_minimal())
```

# References
<div id="refs"></div>
